Laplacian envelope

### 2020-10-22 Notes/laplace.tex
numerics on sqrt bound

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{?.tex} with its contents
%       or include ?.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage[T1]{fontenc}
\usepackage{a4wide}
\usepackage{stix}
% Preamble:
%\usepackage{a4wide}
%\usepackage{ams} for finding documentation only
\usepackage[centertags]{amsmath}
% \usepackage{amssymb}
\usepackage{xcolor}
\usepackage[sort&compress,numbers]{natbib}
\usepackage{ifpdf}
%
\ifpdf
 \usepackage[pdftex,bookmarks,bookmarksopen,pdfstartview=FitH,bookmarksnumbered]{hyperref}
\else
 \usepackage[hypertex]{hyperref}
\fi
% ----------------------------------------------------------------
\vfuzz2pt % Don't report over-full v-boxes if over-edge is small
\hfuzz2pt % Don't report over-full h-boxes if over-edge is small
% ----------------------------------------------------------------
\def\Snospace~{\S{}}
\renewcommand{\sectionautorefname}{\Snospace}
\renewcommand{\subsectionautorefname}{\Snospace}
\renewcommand{\subsubsectionautorefname}{\Snospace}
\usepackage{cleveref}
\usepackage{mathtools}
%
% >> Only for drafts! <<
\usepackage[notref,notcite]{showkeys}
%
\input{sl_graphics_preamble.tex}
\graphicspath{{Figures/}}
%
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%\renewcommand{\baselinestretch}{1.5}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MM}{\mathbf{Q}}
%prob distribution
\newcommand{\pr}{\mathbf{p}}
%equilibrium distribution
\newcommand{\eq}{\pib}
%equilibrium metric
\newcommand{\Eq}{\boldsymbol{\Pi}}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fund}{\mathbf{Z}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\arow}{\boldsymbol{\xi}}
\newcommand{\Lb}{\boldsymbol{\Lambda}}
\newcommand{\lb}{\boldsymbol{\lambda}}
\newcommand{\etw}{\boldsymbol{\eta}^w}
\newcommand{\thb}{\boldsymbol{\theta}}
\newcommand{\omb}{\boldsymbol{\omega}}
\newcommand{\alb}{\boldsymbol{\alpha}}
%synapse models
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\enc}{\mathbf{K}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\F}{\boldsymbol{\Phi}}
%superscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
%with activity indep processes
\renewcommand{\hom}{\mathbf{H}}
\newcommand{\Mh}{\widetilde{\M}}
\newcommand{\frgh}{\widetilde{\W}^{\mathrm{F}}}
\newcommand{\ench}{\widetilde{\enc}}
%eigenvectors
\newcommand{\evr}{\mathbf{u}}
\newcommand{\evl}{\boldsymbol{\eta}}
%snr curves etc
\newcommand{\syn}{\vec{w}}
\newcommand{\synid}{\syn_\text{id}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
\DeclareMathOperator{\NNR}{NNR}
\DeclareMathOperator{\nnr}{NNR}
\newcommand{\snrb}{\overline{\snr}}
\DeclareMathOperator{\env}{Env}
\newcommand{\rh}{\hat{r}}
\newcommand{\CA}{\mathcal{A}}
\newcommand{\CB}{\mathcal{B}}
%avoid double superscripts
\renewcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
% shifted process
\newcommand{\shift}[1]{\widehat{#1}}

% ----------------------------------------------------------------
\input{sl_theorems_preamble.tex}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{Laplacian envelope}
%
% Author List:
%
\author{Subhaneil Lahiri
%
}

@@

When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices $\M^xn  $ whose $i,j$ elements are the transition probabilities from state $i$ to state $j$, 
and \(\nu \) labels the diffrerent types of plasticity event.
Potentiating events have \(\nu=\)pot, depressing events have \(\nu=\)dep.
We also use the notation
%
\begin{equation*}
  (-1)^\nu = 
    \begin{cases}
       1, & \nu = \text{pot}, \\
      -1, & \nu = \text{dep}.
    \end{cases}
\end{equation*}
%
The matrices $\M^\nu$ are transition matrices of discrete time Markov processes.
Their elements must satisfy the following constraints
%
\begin{equation}\label{eq:constr}
  \M^\nu_{ij} \geq 0 
    \quad \forall \; \nu, i, j, \qquad
  \sum_j \M^\nu_{ij} = 1
    \quad \forall \; \nu, i.
\end{equation}
%
The upper bounds $\M^\nu_{ij} \leq 1$ follow automatically from these.

If we treat the off diagonal elements as the independent degrees of freedom, with the off diagonal elements determined by the row sum constraints (second column of \eqref{eq:constr}), the constraints are
%
\begin{equation}\label{eq:constri}
  \M^\nu_{i j} \geq 0
    \quad \forall \; \nu, i, j : i \neq j, \qquad
  \sum_{j : j\neq i} \M^\nu_{ij} \leq 1
    \quad \forall \; \nu, i.
\end{equation}
%
The second of these will be referred to as the ``diagonal'' constraint.

Given a sequence of plasticity events at times $t_1,t_2,t_3,\ldots$, the probability distribution across the internal states at time $t_n$, 
mariginalising over states at earlier times, is
%
\begin{equation}\label{eq:plastseq}
  \pr(t_n) = \pr(t_0)  \M\pot \M\dep \M\dep \M\pot \M\dep \M\pot \M\pot \M\dep \ldots \M\pot .
\end{equation}
%
If the probability of a given plasticity event being potentiating or depressing is $f\pot$ or $f\dep$ respectively, the probability of the sequence above (conditioned on the number of events) is given by $f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot$.

In computing the expectation in \eqref{eq:overlapone} the plasticity event that occurs at $t=0$ is fixed, but all of the previous and subsequent events are marginalised over.
If we marginalise over the types (potentiating/depressing) of these other events (conditioned on the number of events) we find
%
\begin{multline}\label{eq:plastseqtypes}
  \pr(t_n) = \pr(t_0) (f\pot\M\pot+f\dep\M\dep)  \ldots \\
    \ldots (f\pot\M\pot+f\dep\M\dep) 
    \M\potdep (f\pot\M\pot+f\dep\M\dep) \ldots \\
    \ldots (f\pot\M\pot+f\dep\M\dep),
\end{multline}
% 
where the factor of $\M\potdep$ is the event at $t=0$.
If we expand out all of the parenthetical factors, we will get a sum of terms of the form \eqref{eq:plastseq} weighted by their probabilities.

If $t_0 \ll 0$, the effect of the large number of earlier events will be to put the synapse in the steady state distribution, $\eq$, of the stochastic process $(f\pot\M\pot+f\dep\M\dep)$:
%
\begin{equation}\label{eq:plastseqeq}
  \pr(t_n) = \eq \M\potdep (f\pot\M\pot+f\dep\M\dep)^n,
\end{equation}
%
where $n$ is the number of plasticity events that took place.

If the plasticity events occur at Poisson rate $r$, the probability of $n$ events occurring in time $t$ is given by a Poisson distribution with mean $rt$.
If we marginalise over $n$ as well
%
\begin{equation}\label{eq:plastseqav}
\begin{aligned}
  P(\text{state}=i,t \mid \text{pot/dep},0)
    &= \sum_n \frac{(rt)^n}{n!} \e^{-rt} \eq \M\potdep (f\pot\M\pot+f\dep\M\dep)^n \\
    &= \brk{ \eq \M\potdep\, \e^{rt(f\pot\M\pot+f\dep\M\dep-\I) } }_i,
\end{aligned}
\end{equation}
%
where $\eq$ is the steady state distribution of the continuous time transition matrix that appears in the exponential:
%
\begin{equation}\label{eq:eq}
  \eq\frg = 0,
  \qquad\text{where}\quad
  \frg = \sum_\nu f^\nu \M^\nu - \I
    = f\pot\M\pot + f\dep\M\dep - \I.
\end{equation}
%

This results in
%
\begin{equation}\label{eq:overlapwt}
\begin{aligned}
  \av{\synid^1 \syn^1(t)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \e^{rt\frg} \w, \\
  \av{\synid^1 \syn^1(\infty)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \onev\eq \w \\
         &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \e^{rt\frg} \w .
\end{aligned}
\end{equation}
%
Combining these allows us to write the numerator of \eqref{eq:SNRdef} as
%
\begin{equation}\label{eq:signal}
\begin{aligned}
  \av{\synid \cdt \syn(t)} - \av{\synid \cdt \syn(\infty)}
    &= N \eq \enc \, \e^{rt \frg} \w , \\
    \text{where} \qquad
  \enc &= \sum_\nu (-1)^\nu f^\nu (\M^\nu - \I)
    = f\pot(\M\pot-\I)-f\dep(\M\dep-\I)
\end{aligned}
\end{equation}
%
Combining with \eqref{eq:noise} gives
%
\begin{equation}\label{eq:SNRcurveExact}
  \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{rt\frg}\w}
                 {\sqrt{1-(f\pot-f\dep)^2(\eq\w)^2}}.
\end{equation}
%
The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with $f\pot=f\dep$ or $\eq_+=\eq_-$.
We can perform the maximisation in two steps.
First maximise the numerator at fixed $\eq\w$.
Then maximise the ratio \wrt $\eq\w$.
This will allow us to ignore the denominator until the very end.

We can then treat the off-diagonal elements of $\M^\nu$ as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
The diagonal elements are determined by setting the row sums to 1.
We then have
%
\begin{equation}\label{eq:derivpd}
  \pdiff{\M^\mu_{ij}}{\M^\nu_{mn}} 
      = \delta_{\mu\nu}\, \delta_{im}(\delta_{jn}-\delta_{jm}),
\end{equation}
%
It will also be convenient to define the differential operators
%
\begin{equation}\label{eq:pertfe}
  \pdiff{}{\frg_{ij}} = \frac{1}{2f\pot} \pdiff{}{\M\pot_{ij}} + \frac{1}{2f\dep} \pdiff{}{\M\dep_{ij}},
  \qquad
  \pdiff{}{\enc_{ij}} = \frac{1}{2f\pot} \pdiff{}{\M\pot_{ij}} - \frac{1}{2f\dep} \pdiff{}{\M\dep_{ij}},
\end{equation}
%
these satisfy
%
\begin{equation}\label{eq:derivfe}
  \pdiff{\frg_{ij}}{\frg_{mn}} = \pdiff{\enc_{ij}}{\enc_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
  \qquad
  \pdiff{\frg_{ij}}{\enc_{mn}} = \pdiff{\enc_{ij}}{\frg_{mn}} = 0.
\end{equation}
%

If we wish to use the off-diagonal elements of $\frg$ and $\enc$ as the degrees of freedom, the constraints are
%
\begin{equation}\label{eq:constrfe}
\begin{aligned}
  \frg_{i \neq j}   &\geq \abs{ \enc_{i \neq j} }, &\qquad
  \sum_{j\neq i}\frg_{ij} + (-1)^\mu \enc_{ij}  &\leq f^\mu, \\
  \sum_j \frg_{ij} &=0, &
  \sum_j \enc_{ij} &=0.
\end{aligned}
\end{equation}
%

@@

In analogy to the generalised fundamental matrix of a Markov chain \cite{Kemeny1981fund}, define
%
\begin{equation}\label{eq:lfund}
  \fund(s) = (s \I + \onev \arow -\MM)\inv,
\end{equation}
%
where $\arow$ is an arbitrary row vector satisfying $\arow \onev = \frac{1}{\tau_0} \neq 0$.
This reduces to the fundamental matrix at $s=0$.
We can see that $\onev$ is an eigenvector:
%
\begin{equation}\label{eq:lfundrowsum}
  \fund(s)\onev = \frac{\tau_0}{1+s\tau_0}\onev.
\end{equation}
%
Suppose we have some row vector $\mathbf{a}$ such that $\mathbf{a}\onev=0$.
Then $\mathbf{a}\fund(s)$ is independent of $\arow$:
%
\begin{equation}\label{eq:alfund}
  \pdiff{(\mathbf{a}\fund(s))_i}{\arow_k} = -(\mathbf{a}\fund(s)\onev) \fund_{ki}(s)
      = -\frac{\tau_0}{1+s\tau_0} (\mathbf{a}\onev) \fund_{ki}(s) = 0.
\end{equation}
%

Then we also define
%
\begin{equation}\label{eq:lfptb}
  \fptb(s) = (\onem \fund\dg(s)-\fund(s))\D,
  \qquad
  \fptb_{ij}(s) = \frac{\fund_{jj}(s)-\fund_{ij}(s)}{\eq_j}.
\end{equation}
%
This reduces to the mean first-passage time matrix at $s=0$.

@@

This means that the \( \fptb(s) \) have all the same properties as in the case $s=0$.
I particular, we see that the quantity
%
\begin{equation}\label{eq:lkemeny}
  \eta(s) = \sum_j \fptb_{ij}(s)\eq_j
\end{equation}
%
is independent of the starting state $i$, but now $\eta(s) = \tr\fund(s) - \frac{\tau_0}{1 + s \tau_0}$.
We can then define
%
\begin{equation}\label{eq:lpmixtime}
  \eta^{\pm}_i(s) = \sum_j \fptb_{ij}(s) \eq_j \prn{\frac{1 \pm \w_j}{2}} 
      =  \sum_{j\in \pm} \fptb_{ij}(s) \eq_j.
\end{equation}
%
We can arrange the states in order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.
It will be more convenient to use the quantities
%
\begin{equation}\label{eq:lwpmixtime}
  \etw_i(s) = \sum_j \fptb_{ij}(s)\eq_j \w_j
    = \eta^+_i(s) - \eta^-_i(s) 
    = 2\eta^+_i(s) - \eta(s) 
    = \eta(s) - 2\eta^-_i(s) .
\end{equation}
%
Arranging the states in order of decreasing $\etw_i(s)$ is the same as the order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.
The $\etw_i(s)$ can still be used when $\w$ takes more than two values.

@@

The initial SNR is:
%
\begin{equation}\label{eq:initdef}
  \snr(0) = \snrb(0) = \lim_{s\to\infty} sA(s).
\end{equation}
%

@@

The area under the SNR curve is:
%
\begin{equation}\label{eq:areadef}
  \intd[_0^\infty]{t} \snr(t) = \lim_{\tau\to\infty} \tau \snrb(\tau) = A(0).
\end{equation}
%

>>>>

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{?.tex} with its contents
%       or include ?.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage[T1]{fontenc}
\usepackage{a4wide}
\usepackage[tables,graphics,draft]{sl-preamble}
% \usepackage{stix}
%\usepackage{amsldoc} for finding documentation only
\def\Snospace~{\S{}}
\renewcommand{\sectionautorefname}{\Snospace}
\renewcommand{\subsectionautorefname}{\Snospace}
\renewcommand{\subsubsectionautorefname}{\Snospace}
%
\graphicspath{{Figures/}}
%
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%================================================================================
\makeatletter
\DeclareFontEncoding{LS2}{}{\noaccents@}
\makeatother

\DeclareFontSubstitution{LS2}{stix}{m}{n}
\DeclareSymbolFont{xlargesymbols}{LS2}{stixex}{m}{n}
\DeclareMathDelimiter{\lAngle}{\mathopen}{xlargesymbols}{"EC}{xlargesymbols}{"12}
\DeclareMathDelimiter{\rAngle}{\mathclose}{xlargesymbols}{"ED}{largesymbols}{"13}
\DeclareMathDelimiter{\lVvert}{\mathopen}{xlargesymbols}{"F2}{xlargesymbols}{"F5}
\DeclareMathDelimiter{\rVvert}{\mathclose}{xlargesymbols}{"F2}{xlargesymbols}{"F5}

\newcommand{\aav}[1]{\mathopen{}\left\lAngle #1 \right\rAngle\mathclose{}}
\newcommand{\nnrm}[1]{\mathopen{}\left\lVvert #1 \right\rVvert\mathclose{}}
%================================================================================
%
\usepackage{adjustbox}
\newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
\newcommand{\alignmid}[1]{\adjustbox{valign=c}{#1}}
%
\usepackage{paralist}
\newenvironment{myenuma}{\begin{inparaenum}[(a)]}{\end{inparaenum}}
\newenvironment{myenumi}{\begin{inparaenum}[i)]}{\end{inparaenum}}
\newenvironment{myenumA}{\begin{inparaenum}[\bfseries A.]}{\end{inparaenum}}
\newenvironment{myenumI}{\begin{inparaenum}[\bfseries I.]}{\end{inparaenum}}
\newenvironment{myenumr}{\begin{inparaenum}[1.]}{\end{inparaenum}}
%for nested lists, \ref only produces innermost item:  (hacky)
\makeatletter % for internal macros with @
\renewcommand\p@enumii{}
\makeatother
%
%================================================================================
%matrices
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MMm}{Q}
\newcommand{\MM}{\mathbf{\MMm}}
%prob distribution
\newcommand{\prm}{p}
\newcommand{\pr}{\mathbf{\prm}}
%equilibrium distribution
\newcommand{\eqm}{\pi}
\newcommand{\eq}{\boldsymbol{\eqm}}
%equilibrium metric
\newcommand{\Eqm}{\Pi}
\newcommand{\Eq}{\boldsymbol{\Eqm}}
%first passage times
\newcommand{\fptm}{T}
\newcommand{\fpt}{\mathbf{\fptm}}
%off-diag first passage times
\newcommand{\fptbm}{\overline{\fptm}}
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fundm}{Z}
\newcommand{\fund}{\mathbf{\fundm}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\arowm}{\xi}
\newcommand{\arow}{\boldsymbol{\arowm}}
\newcommand{\Lbm}{\Lambda}
\newcommand{\Lb}{\boldsymbol{\Lbm}}
\newcommand{\lbm}{\lambda}
\newcommand{\lb}{\boldsymbol{\lbm}}
\newcommand{\etwm}{\eta^w}
\newcommand{\etw}{\boldsymbol{\eta}^w}
\newcommand{\thbm}{\theta}
\newcommand{\thb}{\boldsymbol{\thbm}}
\newcommand{\ombm}{\omega}
\newcommand{\omb}{\boldsymbol{\ombm}}
\newcommand{\albm}{\alpha}
\newcommand{\alb}{\boldsymbol{\albm}}
%synapse models
\newcommand{\wm}{w}
\newcommand{\w}{\mathbf{\wm}}
\newcommand{\Wm}{W}
\newcommand{\W}{\mathbf{\Wm}}
\newcommand{\MMdm}{M}
\newcommand{\MMd}{\mathbf{\MMdm}}
\newcommand{\encm}{K}
\newcommand{\enc}{\mathbf{\encm}}
\newcommand{\frgm}{\Wm^{\mathrm{F}}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\Fm}{\Phi}
\newcommand{\F}{\boldsymbol{\Fm}}
%eigenvectors
\newcommand{\evrm}{u}
\newcommand{\evr}{\mathbf{\evrm}}
\newcommand{\evlm}{\eta}
\newcommand{\evl}{\boldsymbol{\evlm}}
%superscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
%snr curves etc
\newcommand{\syn}{\vec{w}}
\newcommand{\synid}{\syn_\text{id}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
\DeclareMathOperator{\NNR}{NNR}
\DeclareMathOperator{\nnr}{NNR}
\newcommand{\snrb}{\overline{\snr}}
\DeclareMathOperator{\env}{Env}
% \newcommand{\CA}{\mathcal{A}}
% \newcommand{\CB}{\mathcal{B}}
%avoid double superscripts
\renewcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
% shifted process
\newcommand{\shift}[1]{\widehat{#1}}

% ----------------------------------------------------------------
% \input{sl_theorems_preamble.tex}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{Laplacian envelope}
%
% Author List:
%
\author{Subhaneil Lahiri
%
}

@@

When the synapse is subjected to a plasticity event the internal state will change stochastically, described by the matrices $\MMd^\nu$ whose $i,j$ elements are the transition probabilities from state $i$ to state $j$, 
and \(\nu \) labels the diffrerent types of plasticity event.
Potentiating events have \(\nu=\) pot, depressing events have \(\nu=\) dep.
We also use the notation
%
\begin{equation*}
  (-1)^\nu = 
    \begin{cases}
       1, & \nu = \text{pot}, \\
      -1, & \nu = \text{dep}.
    \end{cases}
\end{equation*}
%
The matrices $\MMd^\nu$ are transition matrices of discrete time Markov processes.
Their elements must satisfy the following constraints
%
\begin{equation}\label{eq:constr}
  \MMdm^\nu_{ij} \geq 0 
    \quad \forall \; \nu, i, j, \qquad
  \sum_j \MMdm^\nu_{ij} = 1
    \quad \forall \; \nu, i.
\end{equation}
%
The upper bounds $\MMdm^\nu_{ij} \leq 1$ follow automatically from these.

If we treat the off diagonal elements as the independent degrees of freedom, with the off diagonal elements determined by the row sum constraints (second column of \eqref{eq:constr}), the constraints are
%
\begin{equation}\label{eq:constri}
  \MMdm^\nu_{i j} \geq 0
    \quad \forall \; \nu, i, j : i \neq j, \qquad
  \sum_{j : j\neq i} \MMdm^\nu_{ij} \leq 1
    \quad \forall \; \nu, i.
\end{equation}
%
The second of these will be referred to as the ``diagonal'' constraint.

Given a sequence of plasticity events at times $t_1,t_2,t_3,\ldots$, the probability distribution across the internal states at time $t_n$, 
mariginalising over states at earlier times, is
%
\begin{equation}\label{eq:plastseq}
  \pr(t_n) = \pr(t_0)  \MMd \pot \MMd \dep \MMd \dep \MMd \pot \MMd \dep \MMd \pot \MMd \pot \MMd \dep \ldots \MMd \pot .
\end{equation}
%
If the probability of a given plasticity event being potentiating or depressing is $f\pot$ or $f\dep$ respectively, the probability of the sequence above (conditioned on the number of events) is given by $f\pot f\dep f\dep f\pot f\dep f\pot f\pot f\dep \ldots f\pot$.

In computing the expectation in \eqref{eq:overlapone} the plasticity event that occurs at $t=0$ is fixed, but all of the previous and subsequent events are marginalised over.
If we marginalise over the types (potentiating/depressing) of these other events (conditioned on the number of events) we find
%
\begin{multline}\label{eq:plastseqtypes}
  \pr(t_n) = \pr(t_0) (f\pot \MMd \pot + f\dep \MMd \dep)  \ldots \\
    \ldots (f\pot \MMd \pot + f\dep \MMd \dep) 
     \MMd \potdep (f\pot \MMd \pot + f\dep \MMd \dep) \ldots \\
    \ldots (f\pot \MMd \pot + f\dep \MMd \dep),
\end{multline}
% 
where the factor of $ \MMd \potdep$ is the event at $t=0$.
If we expand out all of the parenthetical factors, we will get a sum of terms of the form \eqref{eq:plastseq} weighted by their probabilities.

If $t_0 \ll 0$, the effect of the large number of earlier events will be to put the synapse in the steady state distribution, $\eq$, of the stochastic process $(f\pot \MMd \pot+f\dep \MMd \dep)$:
%
\begin{equation}\label{eq:plastseqeq}
  \pr(t_n) = \eq  \MMd \potdep (f\pot \MMd \pot+f\dep \MMd \dep)^n,
\end{equation}
%
where $n$ is the number of plasticity events that took place.

If the plasticity events occur at Poisson rate $r$, the probability of $n$ events occurring in time $t$ is given by a Poisson distribution with mean $rt$.
If we marginalise over $n$ as well
%
\begin{equation}\label{eq:plastseqav}
\begin{aligned}
  P(\text{state}=i,t \mid \text{pot/dep},0)
    &= \sum_n \frac{(rt)^n}{n!} \e^{-rt} \eq  \MMd \potdep (f\pot \MMd \pot+f\dep \MMd \dep)^n \\
    &= \brk{ \eq  \MMd \potdep\, \e^{rt(f\pot \MMd \pot+f\dep \MMd \dep-\I) } }_i,
\end{aligned}
\end{equation}
%
where $\eq$ is the steady state distribution of the continuous time transition matrix that appears in the exponential:
%
\begin{equation}\label{eq:eq}
  \eq\frg = 0,
  \qquad\text{where}\quad
  \frg = \sum_\nu f^\nu \MMd^\nu - \I
    = f\pot \MMd \pot + f\dep \MMd \dep - \I.
\end{equation}
%

This results in
%
\begin{equation}\label{eq:overlapwt}
\begin{aligned}
  \av{\synid^1 \syn^1(t)} &= \eq (f\pot \MMd \pot-f\dep \MMd \dep)\, \e^{rt\frg} \w, \\
  \av{\synid^1 \syn^1(\infty)} &= \eq (f\pot \MMd \pot-f\dep \MMd \dep)\, \onev\eq \w \\
         &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \e^{rt\frg} \w .
\end{aligned}
\end{equation}
%
Combining these allows us to write the numerator of \eqref{eq:SNRdef} as
%
\begin{equation}\label{eq:signal}
\begin{aligned}
  \av{\synid \cdt \syn(t)} - \av{\synid \cdt \syn(\infty)}
    &= N \eq \enc \, \e^{rt \frg} \w , \\
    \text{where} \qquad
  \enc &= \sum_\nu (-1)^\nu f^\nu (\MMd^\nu - \I)
    = f\pot(\MMd \pot-\I)-f\dep(\MMd \dep-\I)
\end{aligned}
\end{equation}
%
Combining with \eqref{eq:noise} gives
%
\begin{equation}\label{eq:SNRcurveExact}
  \SNR(t) = \frac{\sqrt{N} \eq \enc\,\e^{rt \frg}\w}
                 {\sqrt{1-(f\pot-f\dep)^2(\eq \w)^2}}.
\end{equation}
%
The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with $f\pot=f\dep$ or $\eq_+=\eq_-$.
We can perform the maximisation in two steps.
First maximise the numerator at fixed $\eq \w$.
Then maximise the ratio \wrt $\eq\w$.
This will allow us to ignore the denominator until the very end.

We can then treat the off-diagonal elements of $\MMd^\nu$ as the independent degrees of freedom, subject to these constraints, when maximising the SNR.
The diagonal elements are determined by setting the row sums to 1.
We then have
%
\begin{equation}\label{eq:derivpd}
  \pdiff{\MMdm^\mu_{ij}}{\MMdm^\nu_{mn}} 
      = \delta_{\mu\nu}\, \delta_{im}(\delta_{jn}-\delta_{jm}),
\end{equation}
%
It will also be convenient to define the differential operators
%
\begin{equation}\label{eq:pertfe}
  \pdiff{}{\frgm_{ij}} = \frac{1}{2f\pot} \pdiff{}{\MMdm \pot_{ij}} + \frac{1}{2f\dep} \pdiff{}{\MMd \dep_{ij}},
  \qquad
  \pdiff{}{\encm_{ij}} = \frac{1}{2f\pot} \pdiff{}{\MMdm \pot_{ij}} - \frac{1}{2f\dep} \pdiff{}{\MMdm \dep_{ij}},
\end{equation}
%
these satisfy
%
\begin{equation}\label{eq:derivfe}
  \pdiff{\frgm_{ij}}{\frgm_{mn}} = \pdiff{\encm_{ij}}{\encm_{mn}} = \delta_{im}(\delta_{jn}-\delta_{jm}),
  \qquad
  \pdiff{\frgm_{ij}}{\encm_{mn}} = \pdiff{\encm_{ij}}{\frgm_{mn}} = 0.
\end{equation}
%

If we wish to use the off-diagonal elements of $\frg$ and $\enc$ as the degrees of freedom, the constraints are
%
\begin{equation}\label{eq:constrfe}
\begin{aligned}
  \frgm_{i \neq j}   &\geq \abs{ \encm_{i \neq j} }, &\qquad
  \sum_{j\neq i}\frgm_{ij} + (-1)^\mu \encm_{ij}  &\leq f^\mu, \\
  \sum_j \frgm_{ij} &=0, &
  \sum_j \encm_{ij} &=0.
\end{aligned}
\end{equation}
%

@@

In analogy to the generalised fundamental matrix of a Markov chain \cite{Kemeny1981fund}, define
%
\begin{equation}\label{eq:lfund}
  \fund(s) = (s \I + \onev \arow -\MM)\inv,
\end{equation}
%
where $\arow$ is an arbitrary row vector satisfying $\arow \onev = \frac{1}{\tau_0} \neq 0$.
This reduces to the fundamental matrix at $s=0$.
We can see that $\onev$ is an eigenvector:
%
\begin{equation}\label{eq:lfundrowsum}
  \fund(s)\onev = \frac{\tau_0}{1+s\tau_0}\onev.
\end{equation}
%
Suppose we have some row vector $\mathbf{a}$ such that $\mathbf{a}\onev=0$.
Then $\mathbf{a}\fund(s)$ is independent of $\arow$:
%
\begin{equation}\label{eq:alfund}
  \pdiff{(\mathbf{a}\fund(s))_i}{\arowm_k} = -(\mathbf{a}\fund(s)\onev) \fundm_{ki}(s)
      = -\frac{\tau_0}{1+s\tau_0} (\mathbf{a}\onev) \fundm_{ki}(s) = 0.
\end{equation}
%

Then we also define
%
\begin{equation}\label{eq:lfptb}
  \fptb(s) = (\onem \fund\dg(s)-\fund(s))\Eq\inv,
  \qquad
  \fptbm_{ij}(s) = \frac{\fundm_{jj}(s)-\fundm_{ij}(s)}{\eqm_j}.
\end{equation}
%
This reduces to the mean first-passage time matrix at $s=0$.

@@

This means that the \( \fptb(s) \) have all the same properties as in the case $s=0$.
In particular, we see that the quantity
%
\begin{equation}\label{eq:lkemeny}
  \eta(s) = \sum_j \fptbm_{ij}(s) \eqm_j
\end{equation}
%
is independent of the starting state $i$, but now $\eta(s) = \tr\fund(s) - \frac{\tau_0}{1 + s \tau_0}$.
We can then define
%
\begin{equation}\label{eq:lpmixtime}
  \eta^{\pm}_i(s) = \sum_j \fptbm_{ij}(s) \eqm_j \prn{\frac{1 \pm \wm_j}{2}} 
      =  \sum_{j\in \pm} \fptbm_{ij}(s) \eqm_j.
\end{equation}
%
We can arrange the states in order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.
It will be more convenient to use the quantities
%
\begin{equation}\label{eq:lwpmixtime}
  \etwm_i(s) = \sum_j \fptbm_{ij}(s)\eqm_j \wm_j
    = \eta^+_i(s) - \eta^-_i(s) 
    = 2\eta^+_i(s) - \eta(s) 
    = \eta(s) - 2\eta^-_i(s) .
\end{equation}
%
Arranging the states in order of decreasing $\etwm_i(s)$ is the same as the order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.
The $\etwm_i(s)$ can still be used when $\w$ takes more than two values.

@@

The initial SNR is:
%
\begin{equation}\label{eq:initdef}
  \snr(0) = \snrb(0) = \lim_{s\to\infty} sA(s)
      = \sqrt{N} \sum_a \CI_a.
\end{equation}
%

@@

The area under the SNR curve is:
%
\begin{equation}\label{eq:areadef}
  \intd[_0^\infty]{t} \snr(t) = \lim_{\tau\to\infty} \tau \snrb(\tau) 
      = A(0) = \sqrt{N} \sum_a \CI_a \tau_a.
\end{equation}
%
